\mychapter{3}{Lesson 3} %181003

\section{Randomness Extraction}

\textbf{Goal:} extract "real" randomness from "imperfect" randomness.

\bigskip
Suppose to have a biased coin $B s.t. Pr[B=0]=p\neq \frac{1}{2}$ and\\ $Pr[B=1]=1-p$.\\
Can you make it uniform?

\subsection{Von Neumann Extractor}
\begin{enumerate}
    \item Sample $b_1\pickUAR B; b_2\pickUAR B$
    \item If $b_1=b_2$ go to 1.
    \item Else:
    \begin{itemize}
        \item If $b_1=0;b_2=1$ output 1
        \item If $b_1=1;b_2=0$ output 0 
    \end{itemize}
\end{enumerate}
\textbf{Considerations: }
\begin{itemize}
    \item The probability of both cases in 3. is $p(p-1)$
    \item The probability to reach 3. is $2p(p-1)$
    \item $Pr[\text{Fail n-times}]=(1-2p(1-p))^n$ which is very small
\end{itemize}

\textbf{Goal: } Let $X$ be a Random Variable, I want to design an Extractor such that $Ext(X)$ is uniform.
\underline{PROBLEM: } Impossible as stated. What if $X$ is constant? $\implies$ Add the requirement that $X$ has uncertainty (ENTROPY)

\begin{definition}
    MIN-ENTROPY of $X$ (written as $H_{\infty}$) is:

    $$H_{\infty}(X)=-log_2 max_x Pr[X=x]$$
    Where $max_x$ is the x that maximize the probability.
\end{definition}
\begin{example}
    Say $X=\{0,1\}^n$

    \begin{enumerate}
        \item $X$ is constant($Pr[X=\bar{x}]=1, Pr[X=x\neq \bar{x}]=0$)\\
        $\implies H_{\infty}(X)=0$
        \item $X\equiv U_n \implies H_{\infty}(x)=n$ (since $X$ is uniform $Pr[X=x]=2^{-n}$)   
    \end{enumerate}
\end{example}

\textbf{Question: }We want $Ext:\{0,1\}^n \to \{0,1\}$ working $\forall X s.t. H_{\infty}(X)\geq k$\\
\textbf{Claim: } Impossible even for $k=n-1$
\begin{proof}
    Let Ext be any candidate extractor, Let b be any binary value s.t. $|Ext^-1(b)|$ is maximal ($Ext^-1: \{0,1\} \to \{0,1\}^n$)\\
    $\implies |Ext^-1(b)|\geq 2^{n-1}=\frac{2^2}{2}$

    \begin{figure}[ht]
        \centering
        \begin{tikzpicture}[>=latex]
            
            \node (a1) {};
            \node[below=0.3cm of a1] (a2) {};
            \node[below=0.3cm of a2] (a3) {};

            \node[below=0.6cm of a3] (a4) {};
            \node[below=0.3cm of a4] (a5) {};
            \node[below=0.3cm of a5] (a6) {};

            \node[right=4cm of a2] (b1) {$x=0$};
            \node[right=4cm of a5] (b3) {$x=1$};

            \node[ellipse,line width = 1pt, draw=black,minimum size=3cm,fit={(a1) (a6)}] {};

            \node[below=1cm of a6,font=\color{black}\Large] {$\{0,1\}^n$};

            \draw[-,black] (a3) to[out=-10,in=190] (b1.190);
            \draw[-,black] (a2) -- (b1.180);
            \draw[-,black] (a1) to[out=10,in=170] (b1.170);

            \draw[-,black] (a6) to[out=-10,in=190] (b3.190);
            \draw[-,black] (a5) -- (b3.180);
            \draw[-,black] (a4) to[out=10,in=170] (b3.170);
        \end{tikzpicture}
    \end{figure}

    Define $X$ to be uniform over $Ext^-1(b)$ so $H_{\infty}(x)=n-1$ but $Ext(X)=b$ CONSTANT.
    \todo{Help with the proof}
\end{proof}

\begin{theorem}
    \textbf{Way-Out-Seeded-Extractor} allow $X$ to take SMALL uniformly random SEED(catalyst)

    $$Ext: \underbrace{\{0,1\}^d}_{seed(public)}\times \underbrace{\{0,1\}^n}_{input} \to \underbrace{\{0,1\}^m}_{output}$$
\end{theorem}
\begin{definition}
    We say that $Ext:\{0,1\}^d \times \{0,1\}^n \to \{0,1\}^l$ is a $(k,\epsilon)$-extractor ("good") if $\forall x s.t. H_{\infty}(X)\geq k$ we have:
    $$SD((S,Ext(S,x)),(S,U_l))\leq \epsilon$$ 
    where $S\equiv U_d$, $SD$ denotes the statistical distance and we are basically comparing two uniform values ($x$ and $U_l$) 
\end{definition}

\textbf{Idea: } Even if the seed is public the $SD$ must not be influenced. Therefore the two compared variables should be uniform so that $\epsilon$ is small. In this way the output of the extractor can't be distinguished from a uniform string.

\begin{definition}
    \textbf{Statistical Distance:} a distance between two random variables.\\
    Say that $X,X'$ are Random Variables over $\X$.\\
    $SD(x,x')=\frac{1}{2}\sum_{x\in \X }|Pr[X=x]-Pr[X'=x]|$
    \todo{Image of the statistical distance}
\end{definition}

\todo{Leftover hash lemma}